---
title: "PS2 rm test"
output: pdf_document
---
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(readr)
library(tidyverse)
library(dplyr)
library(survey)
library(poliscidata)
gss <- read_csv("gss.csv")
gss_cleaned <- gss %>%
  select(feelings_life,income_family,own_rent,average_hours_worked) %>%
  drop_na(feelings_life) %>%
  mutate(own_rent = coalesce(own_rent, "Don't know"),
         average_hours_worked = coalesce(average_hours_worked, "Don't know"),
         own_rent = ifelse(own_rent == "Owned by you or a member of this household, even if it i...", "Own", 
                           ifelse(own_rent == "Rented, even if no cash rent is paid", "Rent", "Don't know")))

gss_cleaned$income_family <- factor(gss_cleaned$income_family, levels = c("Less than $25,000",
                                                                          "$25,000 to $49,999",
                                                                          "$50,000 to $74,999",
                                                                          "$75,000 to $99,999",
                                                                          "$100,000 to $ 124,999",
                                                                          "$125,000 and more"))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(gss_cleaned, aes(x=factor(feelings_life), fill=own_rent)) + 
  geom_bar() +
  labs(title= "Figure 1: Count of feelings of life categorized by whether they own or rent",
         x = "Feelings of life", y = "Count", fill = "Housing Status")

ggplot(gss_cleaned, aes(x=factor(feelings_life), fill=income_family)) + 
  geom_bar() +
  labs(title= "Figure 2: Count of feelings of life categorized income of the household",
       x = "Feelings of life", y = "Count", fill = "Income of Entire Household")

ggplot(gss_cleaned, aes(x=factor(feelings_life), fill=average_hours_worked)) + 
  geom_bar() +
  labs(title= "Figure 3: Count of feelings of life categorized by average working hours",
       x = "Feelings of life", y = "Count", fill = "Average Working Hours")

```


- figure 1: the majority throughout all the scales of feelings of life own their home, you can see that the proportion of owners is higher in the 7-10 range compared to the rest 1-6 range. the proportion of renters however seem to follow a trend where it increases as feelings of life increase, but it suddenly drop in the 9-10 range.

-figure 2: Seems to be the majority of the Income category for feelings of life in range 7-10 are households with $125 000

-figure 3: In all categroies of feelings of life, the majority of the average working hours is the 30 - 40 hours

```{r, echo=FALSE, message=FALSE, warning=FALSE}

N = 30538825
n = 20331
fpc.srs = rep(N, n)
ucla.design <- svydesign(id=~1, data=gss_cleaned, fpc=fpc.srs)
myregression <- svyglm(feelings_life ~ as.factor(income_family) + as.factor(own_rent) + as.factor(average_hours_worked), 
                  ucla.design)

summary(myregression)

fit.svyglm(myregression)     

summary(gss_cleaned$feelings_life - myregression$fitted.values)
mean((gss_cleaned$feelings_life - myregression$fitted.values)^2)

```


From the coefficients table we know the equation with the base being income of less than
$25000, own/rent status is idk, and avg hours worked = 0 is the following:

$$
FeelingsOfLife = 6.29 + 0.37x_{income1} + 0.54x_{income2} + 0.62x_{income3} 
+ 0.62x_{income4} + 0.62x_{income5} \\
+ 0.45x_{own} + 0.09x_{rent} \\
+ 0.91x_{work1} + 0.87x_{work2} + 0.95x_{work3} + 1.01x_{work4} + 0.98x_{work5}
$$

The adjusted r square is 0.043 suggests we have a very inaccurate model but it could just be from the fact we
have so many predictor variables. So performing an MSE we have 2.59 


```{r, echo=FALSE, message=FALSE, warning=FALSE}

summ <- list("Feelings of Life" = 
               list("Min" = min(gss_cleaned$feelings_life),
                    "Max" = max(gss_cleaned$feelings_life),
                    "Mean" = mean(gss_cleaned$feelings_life),
                    "Median" = median(gss_cleaned$feelings_life)))

x <- data.frame("Stats" = c("Min", "Max", "Mean", "Median"),
                "Numbers" = c(min(gss_cleaned$feelings_life),
                              max(gss_cleaned$feelings_life),
                              mean(gss_cleaned$feelings_life),
                              median(gss_cleaned$feelings_life)))
x

setNames(data.frame(table(gss_cleaned$income_family)), c("Family Income", "Frequency"))
setNames(data.frame(table(gss_cleaned$own_rent)), c("Own/Rent", "Frequency"))
setNames(data.frame(table(gss_cleaned$average_hours_worked)), c("Average Hours Worked", "Frequency"))

```

















